---
title: "Reproducible Research Project 1"
output: md_document
date: "2023-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(tidyverse)
```

Code for reading in the dataset and/or processing the data

```{r}
activity <- read.csv("activity.csv")
```

Histogram of the total number of steps taken each day

```{r}
activitySummarized <- activity %>%
  group_by(date) %>%
  summarize(steps = sum(steps))

ggplot(activitySummarized, aes(x = ymd(date), y = steps)) +
  geom_bar(stat = "identity") +
  xlab("Date") +
  ylab("Steps")
```

Mean and median number of steps taken each day

```{r}
mean(na.omit(activitySummarized$steps))
median(na.omit(activitySummarized$steps))
```

Time series plot of the average number of steps taken

```{r}
plot(ymd(activitySummarized$date), activitySummarized$steps, type = "l", xlab = "Date", ylab = "Steps")
```

The 5-minute interval that, on average, contains the maximum number of steps

```{r}
activitySteps <- activity %>%
  group_by(interval) %>%
  summarize(averageSteps = mean(na.omit(steps)))

activitySteps[which(activitySteps$averageSteps == max(activitySteps$averageSteps)), ]$interval
```

Code to describe and show a strategy for imputing missing data

```{r}
activityImputed <- activity
activityImputed$steps[is.na(activityImputed$steps)] <- mean(na.omit((activity$steps)))
```

Histogram of the total number of steps taken each day after missing values are imputed

```{r}
ggplot(activityImputed, aes(x = ymd(date), y = steps)) +
  geom_bar(stat = "identity") +
  xlab("Date") +
  ylab("Steps")
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
activity$weekend <- if_else(wday(activity$date) %in% c(1, 7), "Weekend", "Weekday")

activityWeekend <- group_split(activity, weekend)[[1]] %>%
  group_by(interval) %>%
  summarize(averageSteps = mean(na.omit(steps)))
activityWeekend$weekend <- "Weekend"

activityWeekdays <- group_split(activity, weekend)[[2]] %>%
  group_by(interval) %>%
  summarize(averageSteps = mean(na.omit(steps)))
activityWeekdays$weekend <- "Weekday"

activityWeek <- rbind(activityWeekend, activityWeekdays)

ggplot(activityWeek, aes(x = interval, y = averageSteps)) +
  geom_bar(stat = "identity") +
  facet_wrap(~weekend) +
  xlab("Interval") +
  ylab("Steps")
```